## Section 0
# Install packages
```{r, slide 17}
install.packages('readr')
install.packages('readxl')
install.packages('tidyr')
install.packages('plyr')
install.packages('dplyr')
install.packages("memisc")
install.packages("car")
install.packages("hunspell")
install.packages("ggplot2")
```
# Library packages
```{r, library packages, slide 18}
library(readr) # Read Tabular data, such as CSV files
library(readxl) #Read excel files
library(tidyr) #Easily Tidy data with 'spread()' and 'gather()' Functions
library(plyr) #Tools for splitting, Applying and Combining data
library(dplyr) # A grammar of data manipulation
```
# Import dataset
```{r, slide 19}
## Method 1 By setting workind directory 
# "C:/Users/flyff/Desktop/UMD/Job-Career/Tesla Gov"
setwd("PASTE FILE FOLDER PATH HERE")
getwd()                                     
ACLED_df<- read.csv("ACLED data-2018-04-11-Israel-Palestine.csv")

## Method 2 By inputting the full dirctory path
ACLED_dftry2<- read.csv("C:/Users/flyff/Desktop/UMD/Job-Career/Tesla Gov/ACLED data-2018-04-11-Israel-Palestine.csv")

identical(ACLED_df,ACLED_dftry2) #This tells you if the two data frames are identifical
rm(ACLED_dftry2) # Removes whatever object you put in it
```
## Exercise 0
```{r, slide 20}
x <- c(1:10, 1:5, NA)
x
length(x)                   # length of x 
table(x)                    # Frequency counts of entries
max(x, na.rm = TRUE)        # Find the maximum value in x, excluding missing values
min(x, na.rm = TRUE)        # minimum
mean(x, na.rm = TRUE)       # mean
median(x, na.rm = T)        # median 
sum(x, na.rm = T)           # sum
var(x, na.rm = T)           # variance
sd(x, na.rm = T)            # standard deviation

is.na(x)                    # check if each element in x is missing
which(is.na(x))             # the index of missing value in the vector x
which(x == 1)               # the index of a particular value in the vector x
```
## Section 1
# Examing the dataset
```{r, slide 22}
View(ACLED_df)                # To view data (or just click it in the environment)

dim(ACLED_df)                 # Get dimensions - how many observations of how many variables(or just look at the data in the environment)

head(ACLED_df)                # The top few rows of a variable

nrow(ACLED_df)                # number of rows

ncol(ACLED_df)                # number of columns

colnames(ACLED_df)            # name of columns (the variable names) in dataset  
rownames(ACLED_df)            # row names

str(ACLED_df)                 # Compactly displays the structure of an arbitrary R object

unique(ACLED_df$actor1) #See how unique responses you have within a column

library(memisc) # Library containng the codebook function
#checking for empty responses
ACLEDyear_codebook<-codebook(as.factor(ACLED_df$year)) # Generates a codebook of a dataset

ACLEDyear_codebook # The number and percent of observations that were year 2017 and 2018

codebook(complete.cases(ACLED_df$year)) # Of all the observations present are any empty?
```
# Subsetting data
```{r, slide 23}
# Subsetting by row and/or column
ACLED_df[1,1]                 # first row first column
ACLED_df[,1]                  # first column
ACLED_df[1,]                  # first row
ACLED_df[,1:2]                # first 2 columns
# Subsetting by vatiable 
ACLED_df$fatalities

# subsetting by specific criteria
ACLED_df[which(ACLED_df$fatalities > 0),]
ACLED_df[which(ACLED_df$source == "Agence France Presse"),]
```
# Summary Report
```{r, slide 24}
# Summary Report Method 1
summary(ACLED_df)                  # Summary of each of the variables
# Summary Report Method 2
aggregate(ACLED_df$fatalities, by= list(ACLED_df$event_type), mean)
# The mean of fatalities by event type
```
## Exercise 1
```{r}
# G1 How many fatalities occurred by event type and year?
aggregate(ACLED_df$fatalities, by= list(ACLED_df$year, ACLED_df$event_type), sum)

# G2 Using actor1 how many fatalities were the Unidentified Armed Group (Palestine) involved in?
aggregate(ACLED_df$fatalities, by= list(ACLED_df$actor1), sum)
min(ACLED_df$fatalities)

# G3 Which day had the most fatalities?
aggregate(ACLED_df$fatalities, by= list(ACLED_df$event_date), sum)
```
## Section 2
# Recoding  
```{r, slide 27-29}
# Replace the first column with the character data ID
colnames(ACLED_df)[1]
colnames(ACLED_df)[1]<- "data ID"

# Replace the first row with the character first_row
rownames(ACLED_df)[1]
rownames(ACLED_df)[1]<- "first_row"

# Replace all the data in a field with a number
data$Grade <- 5

# Replace all the data in a field with with text
data$Grade <- "Five"

# Replace all the data in a field with NA (missing data)
data$Grade <- NA

# Replace the data in a field based on equal to some value
data$Grade[data$Grade==5] <- "Grade Five"

# Or replace based on greater than or equal to some value
data$Grade[data$Grade<=5] <- "Grade Five or Less"

# Or replace based on equal to some text
data$Grade[data$Grade=="Five"] <- "Grade Five"

# Or replace only missing data
# Note that ==NA does not work!
data$Grade[is.na(data$Grade)] <- "Missing Grade"
```
#Recoding with the car package
```{r}
# Load the car package
library(car)
# Recode grade 5 to grade 6
Data$Grade<-recode(Data$Grade,"5=6")
# Recode grade 5 to grade 6 based on text
Data$Grade<-recode(Data$Grade,"'Grade Five'=5")
# Recode grade 5 to grade 6
Data$Grade<-recode(Data$Grade,"c(1,2,3,4,5)='Five or Less'")
# Create a new field called NewGrade
Data$NewGrade <- recode(Data$Grade,"5='Elementary'")
# Recode grade 3 to NA
Data$Grade<-recode(Data$Grade,"3=NA")
# Or recode NA to 7
Data$Grade <- recode(Data$Grade,"NA=7")
# Recode grade 5 to grade 6 and grade 6 to grade 7
SchoolData$Grade<-recode(SchoolData$Grade,"5=6;6=7")
# Recode grades 1 through 5 to Elementary
SchoolData$Grade<-recode(SchoolData$Grade,"1:5='Elementary'")
# Recode the lowest grade through 5 to Elementary
SchoolData$Grade<-recode(SchoolData$Grade,"lo:5='Elementary'")

# Recode grade 9 to the highest grade to High School
SchoolData$Grade<-recode(SchoolData$Grade,"9:hi='High School'")
# Recode grades
SchoolData$Grade<-recode(SchoolData$Grade,"1:5='Elementary';6:8='Middle;else='High'")
```

## Sorting dataset
```{r, slide 30}
## Sorting data
data[order(data$variable), ] # would not change the original data
data.sorted <- data[order(data$var1,data$var2), ] # Sort by two vectors, and save the sorted data in a new dataset
Acled_sorted <- ACLED_df[order(ACLED_df$source, ACLED_df$year), ] 
```
## Exercise 2
```{r}
# G1 Rename the column "source" to "opinion"
colnames(ACLED_df)[25]<- "opinion"

# G2 Sort the dataset by "opinion" and "year"

Acled_sorted_Ex2 <- ACLED_df[order(ACLED_df$opinion, ACLED_df$year), ] 
# G3 Recode all 'BBC' observations to 'BritishBeingCool'
Acled_sorted_Ex2$opinion<-recode(Acled_sorted_Ex2$opinion,"'BBC'='BritishBeingCool'")

```
## Section 3
# Mispellings/miscategorizations
```{r, 33}
library(hunspell)

words <- c("beer", "wiskey", "wine")
# Checks individual words
hunspell_check(words)
 
correct <- hunspell_check(words)
correct # Correct words

# Suggests words for the incorrect bunch of words
hunspell_suggest(words[!correct])

# The hunspell_find function takes a character vector with text (in plain, latex or man format) and returns a list with incorrect words for each line.
bad_words <- hunspell_find("spell checkers are not neccessairy for langauge ninja's")
print(bad_words)
# Finally hunspell_suggest is used to suggest correct alternatives for each (incorrect) input word.
hunspell_suggest(bad_words[[1]])

# Arabic example
ar <- dictionary("C:/Users/flyff/AppData/Local/RStudio-Desktop/dictionaries/languages-user/ar.dic")
ArabicCorrect<- "??????????"#Hello
hunspell_check(ArabicHello,dict = ar)
hunspell_suggest(ArabicHello,dict = ar)

ArabicIncorrect <-"????????"#Jams
hunspell_check(ArabicIncorrect,dict = ar)
hunspell_suggest(ArabicIncorrect,dict = ar)
```
## Section 4
# Merging data frames
```{r, 36}
ACLED_df_2 <- read.csv("C:/Users/flyff/Desktop/UMD/Job-Career/Tesla Gov/2018-04-12-2018-08-01-Middle_East-Israel-Palestine.csv")
str(ACLED_df_2)
Merged_ACLED_df<- rbind(ACLED_df,ACLED_df_2)
Merged_ACLED_df
```
# Plotting data
```{r}
library(ggplot2)
# Barplot, Good.
barplot(table(Merged_ACLED_df$event_type), xlab = "Event Type", ylab = "Frequency")
title(main = list("Frequency of Events via Merged ACLED data", font = 4))
# Histogram, Better.
ggplot (ACLED_df, aes(x=fatalities,fill = event_type ))   +   		geom_histogram ()
# Hisogram with more arguments, Best.
ggplot (ACLED_df, aes(x=fatalities,fill = event_type ))   +   		geom_histogram (position = "dodge", color = "white", binwidth = 1) 
# Even better than the tbest
ggplot (ACLED_df[ACLED_df$fatalities<2,], aes(x=fatalities,fill = event_type ))   +   		geom_histogram (position = "dodge", color = "white", binwidth = 1) 
# Plot of the summary data


Merged_ACLED_agg <- aggregate(Merged_ACLED_df$fatalities, by=list(Merged_ACLED_df$event_type, Merged_ACLED_df$country), mean)


ggplot(Merged_ACLED_agg, aes(x=Group.2, y= x, fill = Group.1)) + 
  geom_bar(position = "dodge", stat="identity", color = "white")+
  labs( title = "Fatalities by Area",
        subtitle = " ",
        caption = " ",
        x = "Country",
        y = "Fatalities",
        fill = "Event Type")
```
## Exercise 4
```{r}
# Basic violin plot
p <- ggplot(ACLED_df, aes(x=event_type, y=fatalities)) + 
  geom_violin()
p
# Rotate the violin plot
p + coord_flip()
# Set trim argument to FALSE
ggplot(ACLED_df, aes(x=country, y=fatalities)) + 
  geom_violin(trim=TRUE)+ coord_flip()+ scale_y_discrete(limits=c(.5, 1)) + stat_summary(fun.y=mean, geom="point", shape=23, size=2)
```
## Section 5
## Multiple linear regression
```{r}
# A summary of the multiple linear regression of fatalities on country and event type
summary(lm(fatalities~country + event_type, Merged_ACLED_df))
```
#Country codes
```{r}
# install.packages(countrycode)
library(countrycode) #Convert Country names and Country codes
countrycode(c('USA', 'DZA'), 'iso3c', 'cown')
countrycode('Albania', 'country.name', 'iso3c')
countrycode('Albanien', 'country.name.de', 'iso.name.fr')
```

